2.4 PRNT50

### 2025-04-16 2_scripts/2.4 PRNT50.qmd
Update figures and misc analysis

```{r}
# ggplot
p09.lab<-c("APL-Abbott (COI=402)", "CBS-Roche (COI=179)", 
           "APL-Abbott (COI=50)", "CBS-Roche (COI=0.8)")
p09c<-tb_APLCBS2 %>%
  ggplot(aes(x=samplemonth, y=prop, colour=interaction(study, COI_ty),
                                    linetype=interaction(study, COI_ty))) +
  geom_point() +
  geom_line()  +
  theme_bw() +
  xlab("Time of Sample Collection") + 
  ylab("Proportion")+ 
  scale_colour_manual("", values = c("red", "blue", "red", "blue"),
                      labels = p09.lab)+
  scale_linetype_manual("", values=c(3,3,1,1), labels = p09.lab) 
#  theme(legend.position="top")+
#  guides(color = guide_legend(nrow = 2))

@@

#### The unvaccinated stratum
```{r}
# add proportion of the stratum into the data table
tb_abbsv0<-merge(tb_abbsv0, select(tb_abbsv1, samplemonth, n.all),
      by="samplemonth", all.x=T, all.y=F)
tb_abbsv0<-tb_abbsv0 %>%
  mutate(
    prop.stra = n/n.all
  )
p09v0.lab<-c("APL-Abbott (COI=50)", "APL-Abbott (COI=402)",
             "Proportion of the stratum")

@@

#### Figure panel
```{r}
# use cowplot package
p09.upper<-cowplot::plot_grid(p09c + ylab("Proportion") + xlab ("") +
                           theme(legend.position='top') +
                           guides(color = guide_legend(nrow = 2)), 
                           p09v0 + xlab ("") + ylab(""),
                           labels = c("A", "B"),
                           nrow=1,  align = "h")

@@

png(file="../4_output/figs/Fig.4 Immunity.png", 
    width=2400, height=2000, res=300)
p09all
dev.off()
```

>>>>

```{r}
# ggplot
p09.lab<-c("APL (PRNT-50 cutoff)", "CBS (PRNT-50 cutoff)", 
           "APL (manufacturer cutoff)", "CBS (manufacturer cutoff)")
p09c<-tb_APLCBS2 %>%
  ggplot(aes(x=samplemonth, y=prop, colour=interaction(study, COI_ty),
                                    linetype=interaction(study, COI_ty))) +
  geom_point() +
  geom_line()  +
  theme_bw() +
  xlab("Time of Sample Collection") + 
  ylab("Proportion")+ 
  scale_colour_manual("", values = c("red", "blue", "red", "blue"),
                      labels = p09.lab)+
  scale_linetype_manual("", values=c(3,3,1,1), labels = p09.lab)+
  ggtitle("Unstratified")
#  theme(legend.position="top")+
#  guides(color = guide_legend(nrow = 2))

@@

#### The unvaccinated stratum
```{r}
# add proportion of the stratum into the data table
tb_abbsv0<-merge(tb_abbsv0, select(tb_abbsv1, samplemonth, n.all),
      by="samplemonth", all.x=T, all.y=F)
tb_abbsv0<-tb_abbsv0 %>%
  mutate(
    prop.stra = n/n.all
  )
p09v0.lab<-c("APL (manufacturer cutoff)", "APL (PRNT-50 cutoff)",
             "Proportion of the stratum")

@@

#### Figure panel
```{r}
# use cowplot package
p09.upper<-cowplot::plot_grid(p09c + ylab("Proportion") + xlab ("") +
                           coord_cartesian(xlim=as.yearmon(c("Jan 2021", "Oct 2022"))) +
                           theme(legend.position='top') +
                           guides(color = guide_legend(nrow = 2)), 
                           p09v0 + xlab ("") + ylab(""),
                           labels = c("A", "B"),
                           nrow=1,  align = "h")

@@

png(file="../4_output/figs/Fig.4 Immunity.png", 
    width=2400, height=2000, res=300)
p09all
dev.off()
```
## 5. Additional analysis for manuscript
```{r}
# median age in CBS Alberta and APL
summary(CBSs_AB$age) 
summary(APLs$age) # age=0?

APLs %>%
  .[.$age %in% c(0:17),]

# age categories
CBSs_AB$age_cat[CBSs_AB$age %in% c(16:17)]<-"<18"
CBSs_AB$age_cat[CBSs_AB$age %in% c(18:35)]<-"18-35"
CBSs_AB$age_cat[CBSs_AB$age %in% c(36:64)]<-"36-64"
CBSs_AB$age_cat[CBSs_AB$age %in% c(65:199)]<-"65+"
table(CBSs_AB$age_cat, useNA = "ifany")
prop.table(table(CBSs_AB$age_cat, useNA = "ifany"))

APLs$age_cat[APLs$age %in% c(0:17)]<-"<18"
APLs$age_cat[APLs$age %in% c(18:35)]<-"18-35"
APLs$age_cat[APLs$age %in% c(36:64)]<-"36-64"
APLs$age_cat[APLs$age %in% c(65:199)]<-"65+"
table(APLs$age_cat)
prop.table(table(APLs$age_cat))

# Sex
table(CBSs_AB$sex)
prop.table(table(CBSs_AB$sex))

table(APLs$gender)
prop.table(table(APLs$gender))
```

Among samples from our auxiliary datasets tested with both Abbott and Roche Anti-N assays, the COI was lower with Abbott than with Roche in X% of samples
```{r}
# This involves directly comparison btw Roche anti-N and Abbott-anti-N

# How many samples each?
table(dta_all$study)

# Non missing in Abbott & Roche anti-N
table(dta_all[(!is.na(dta_all$roche_n) & !is.na(dta_all$abbott_n)),]$study)
dta_nmiss<-dta_all[(!is.na(dta_all$roche_n) & !is.na(dta_all$abbott_n)),]

table(dta_nmiss$abbott_n<dta_nmiss$roche_n)
778/(778+43)
```

```{r}
#Numbers at 2022 July
tb_melt_2 %>%
  .[.$samplemonth == as.yearmon("July 2022"),]
```

#### PCR-verified test positivity:
```{r}
# How many test results are previously PCR-verified?

# CBS data (AB subset) - ver Dec 2023, this should be the same ver of CBS data we received by late Nov of 2023, the same ver of data we used for the project. No PCR test results were found.

# How many records in AB?
CBSs %>%
  filter(province=="AB") %>%  # N=124,008
  select(pid, sampledate) %>%
  unique()                    # No duplicated record, good

# APL data: check the APL data ev2, this one is uploaded to the jmy folder on the server
load("/home/jchen/../../jmy/APL/RFD4682_ev2.RData")

# N=214776, de-duplicated, good.

# Is sero-id unique (this is the sample ID)
length(unique(RFD4682_ev2$SERO_ID)) # N=214,776, good
length(unique(RFD4682_ev2$SID)) # N=213,139, duplicates, use SERO_ID

# how many has valid PCR test results?
t1<-table(RFD4682_ev2$NAAT_STATUS, useNA = 'ifany')
t1
prop.table(t1)
```

```{r}
# Unstratified test positivity 
# Abbott Anti-N
RFD4682_ev2 %>%
  filter(NAAT_STATUS == "Positive") %>%
  count(`N-IgG_INTERP`)

# Abbott Anti-S-RBD
RFD4682_ev2 %>%
  filter(NAAT_STATUS == "Positive") %>%
  count(`RBD-IgGII_INTERP`)
```
```{r}
# Stratified analysis
RFD4682_ev2$DAYS_SINCE_NAAT<-as.numeric(RFD4682_ev2$DAYS_SINCE_NAAT)
hist(RFD4682_ev2$DAYS_SINCE_NAAT)

RFD4682_ev2$NAAT_month_cat<-rep(NA, length(RFD4682_ev2$SERO_ID))
RFD4682_ev2[!is.na(RFD4682_ev2$DAYS_SINCE_NAAT) &
            0<=RFD4682_ev2$DAYS_SINCE_NAAT &
            RFD4682_ev2$DAYS_SINCE_NAAT<90,]$NAAT_month_cat<-"<3 months"
RFD4682_ev2[!is.na(RFD4682_ev2$DAYS_SINCE_NAAT) &
            90<=RFD4682_ev2$DAYS_SINCE_NAAT &
            RFD4682_ev2$DAYS_SINCE_NAAT<180,]$NAAT_month_cat<-"3-6 months"
RFD4682_ev2[!is.na(RFD4682_ev2$DAYS_SINCE_NAAT) &
            180<=RFD4682_ev2$DAYS_SINCE_NAAT &
            RFD4682_ev2$DAYS_SINCE_NAAT<360,]$NAAT_month_cat<-"6-12 months"
RFD4682_ev2[!is.na(RFD4682_ev2$DAYS_SINCE_NAAT) &
            360<=RFD4682_ev2$DAYS_SINCE_NAAT,]$NAAT_month_cat<-">12 months"
table(RFD4682_ev2$NAAT_month_cat, useNA = 'ifany')

RFD4682_ev2 %>%
  filter(NAAT_STATUS == "Positive") %>%
  group_by(NAAT_month_cat) %>%
  count(`N-IgG_INTERP`)
```
